---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
requireNamespace("tibble")
library(electionca)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  dpi = 300
)
```

# electionca

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![Travis build status](https://travis-ci.org/paleolimbot/electionca.svg?branch=master)](https://travis-ci.org/paleolimbot/electionca)
<!-- badges: end -->

The goal of electionca is to provide Canadian (general) election data in an easily accessible format for R users. It is based on the Library of Parliament's [ParlInfo site](https://lop.parl.ca/sites/ParlInfo/default/en_CA/), but also includes geography for elections since 2003 ([2003](https://open.canada.ca/data/en/dataset/78400aed-2370-4437-97ca-7563c21bacb1), [2013](https://open.canada.ca/data/en/dataset/10801c67-7f18-4ea1-bda7-8962abfc5578) and [2015](https://open.canada.ca/data/en/dataset/737be5ea-27cf-48a3-91d6-e835f11834b0)). Approximate geography is derived for historical ridings based on riding associations noted by the Library of Parliament.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("paleolimbot/electionca")
```
If you can load the package, you're all set!

```{r}
library(electionca)
```

## Example

This package contains geographic information about [ridings](https://lop.parl.ca/sites/ParlInfo/default/en_CA/ElectionsRidings/Ridings) (`ridings`), [results from each riding](https://lop.parl.ca/sites/ParlInfo/default/en_CA/ElectionsRidings/Elections) for each general election (`results`), and low-resolution boundaries, where these are known (2006-present). It contains convenience functions to plot election results in bar, cartogram, and map format:

```{r test-plot}
plot_votes()
plot_seats()
plot_election()
plot_election_map()
```

The easiest way to access the data in this package is through the `election_results()` function, which modifies the raw data slightly to provide the data that you probably want to work with:

```{r}
election_results()
```

If you're looking to do a more comprehensive analysis, you can access the raw data using `results`, `ridings`, and `boundaries`.

```{r example-data}
results
ridings
boundaries
```

For examples of how to use the raw data to create the plots returned by `plot_election_map()` and `plot_election()`, see `vignette("plot_election", package = "electionca")`.
