---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
library(sf)
library(tidyverse)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# electionca

<!-- badges: start -->
<!-- badges: end -->

The goal of electionca is to provide Canadian (general) election data in an easily accessible format for R users. It is based on the Library of Parliament's [ParlInfo site](https://lop.parl.ca/sites/ParlInfo/default/en_CA/), but also includes geography for elections since 2003 ([2003](https://open.canada.ca/data/en/dataset/78400aed-2370-4437-97ca-7563c21bacb1), [2013](https://open.canada.ca/data/en/dataset/10801c67-7f18-4ea1-bda7-8962abfc5578) and [2015](https://open.canada.ca/data/en/dataset/737be5ea-27cf-48a3-91d6-e835f11834b0)). Approximate geography is derived for historical ridings based on riding associations noted by the Library of Parliament.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("paleolimbot/electionca")
```
## Example

This package contains geographic information about [ridings](https://lop.parl.ca/sites/ParlInfo/default/en_CA/ElectionsRidings/Ridings) (`ridings`), [results from each riding](https://lop.parl.ca/sites/ParlInfo/default/en_CA/ElectionsRidings/Elections) for each general election (`results`), and low-resolution boundaries, where these are known (2004-present).

```{r example}
library(electionca)
results
ridings
boundaries
```
