---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
library(sf)
library(tidyverse)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# electionca

<!-- badges: start -->
<!-- badges: end -->

The goal of electionca is to provide Canadian (general) election data in an easily accessible format for R users. It is based on the Library of Parliament's [History of the Federal Electoral Ridings, 1867-2010](https://open.canada.ca/data/en/dataset/ea8f2c37-90b6-4fee-857e-984d3060184e), but also includes results from the [2011](https://www.elections.ca/content.aspx?section=ele&document=index&dir=pas/41ge&lang=e) and [2015](https://www.elections.ca/content.aspx?section=ele&document=index&dir=pas/42ge&lang=e) general elections. Geography for elections since 2003 was derived from the [2003](https://open.canada.ca/data/en/dataset/78400aed-2370-4437-97ca-7563c21bacb1) and [2015](https://open.canada.ca/data/en/dataset/737be5ea-27cf-48a3-91d6-e835f11834b0) riding boundaries; geography for previous elections is currently missing.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("paleolimbot/electionca")
```
## Example

This package contains geographic information about ridings (`boundaries`), and results from each riding, for each general election (`results`).

```{r example}
library(electionca)
boundaries
results
```
