---
title: "Plotting election data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Plotting election data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library(dplyr)
library(ggplot2)
library(forcats)
library(lubridate)
library(electionca)
theme_set(theme_bw())
requireNamespace("sf")

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

If you're looking to make a simple plot of election winners, you probably want to use `plot_election()` or `plot_election_map()`. If you can't make the plot you want using those, this vignette is for you. We'll use the [tidyverse](https://tidyverse.org/) and the [lubridate](https://lubridate.tidyverse.org/) packages. I happen to think the colours work better with `ggplot2::theme_bw()`, so I'm going to set that as the default theme.

```{r, eval = FALSE}
library(tidyverse)
library(lubridate)
library(electionca)
theme_set(theme_bw())
```

## Raw data and bar charts

The `election_results()` function is designed to give you the data that you probably want to work with. Use the `years` and `provinces` arguments to filter for specific elections and provinces, or use [dplyr](https://dplyr.tidyverse.org/) functions to get the data in the right form.

```{r}
on_2015_2019 <- election_results(2015:2019, "Ontario")
on_2015_2019
```

Data in this form is suitable for aggregation, either by dplyr's `group_by() %>% summarise()` or using some kind of aggregator `stat_*()` in ggplot2. In this case, we're plotting the number of votes received by each party in province of Ontario in the 2015 and 2019 general elections (you could also do this using `plot_votes()`). Note the use of `forcats::fct_lump()` and `forcats::fct_reorder()` to plot the most popular parties (if you don't use these, you'll get a whole lot of parties that received few votes). I also use `scale_fill_party()` here, not because it's necessary, but because it makes the plot much more user-friendly (when is the last time you saw an election result plot without party colours!).

```{r}
on_2015_2019 %>% 
  mutate(
    party = party %>% 
      fct_lump(n = 4, w = votes) %>% 
      fct_reorder(votes, sum)
  ) %>% 
  ggplot(aes(party, fill = party)) +
  stat_count(aes(weight = votes / 1e6)) +
  facet_wrap(vars(year(election_date))) +
  scale_fill_party(guide = "none") +
  coord_flip() +
  labs(x = NULL, y = "Millions of votes")
```

This reflects the number of votes, but not the number of seats. To plot the number of seats, we need to `filter()` the election results data to only the winners of each riding (`result == "Elected"`), and stop scaling by `votes`, since each riding should take up the same amount of space regardless of how many votes the winner received (you could also do this using `plot_seats()`. To order the parties properly, we need to order by frequency rather than by the number of votes that each received, which I do here using `forcats::fct_infreq()` and `forcats::fct_rev()`.

```{r}
on_2015_2019 %>% 
  filter(result == "Elected") %>% 
  mutate(
    party = party %>% 
      fct_infreq() %>% 
      fct_rev()
  ) %>% 
  ggplot(aes(party, fill = party)) +
  stat_count() +
  facet_wrap(vars(year(election_date))) +
  scale_fill_party(guide = "none") +
  coord_flip() +
  labs(x = NULL, y = "Number of seats")
```

Another way to plot these data would be to put the year on the x-axis and stack all the bars on top of eachother to make it easier to compare the results from year to year. I'll use the number of seats here, but one could also do this with the number of votes using the same `stat_count()` as above.

```{r}
on_2015_2019 %>% 
  filter(result == "Elected") %>% 
  mutate(
    party = party %>% 
      fct_infreq() %>% 
      fct_rev()
  ) %>% 
  ggplot(aes(factor(year(election_date)), fill = party)) +
  stat_count() +
  scale_fill_party(guide = "none") +
  labs(x = "Election year", y = "Number of seats")
```

Another plot you might want to make is one comparing how various provinces voted. The code for this is similar to the election year plot above, but instead uses `province` on the x axis. Because or original data was only for Ontario, we need to fetch a new subset of the data using `election_results()`. I think this makes more sense using `votes` rather than seats, so I'm going to use that version of the plot. I also think it makes more sense to use `position = "fill"` here, because we really want to see the percentage of how the provinces voted, rather than the number of votes cast in each province (although both are useful plots!).

```{r}
election_results(year = 2015:2019) %>% 
  mutate(
    party = party %>% 
      fct_lump(n = 5, w = votes) %>% 
      fct_reorder(votes, sum) 
  ) %>% 
  ggplot(aes(x = province, fill = party)) +
  stat_count(aes(weight = votes), position = "fill") +
  scale_fill_party() +
  scale_y_continuous(labels = scales::percent_format()) +
  coord_flip() +
  labs(x = NULL, y = "Proportion of votes")
```

## Incorporating geography

More often than not, you've probably seen an election map with the whole of Canada pictured, most of which are the three northern ridings of Nunavut, Northwest Territories, and Yukon Territory. This grossly overestimates the importance of these ridings, so as an alternative, the electionca package includes some information that can be used to plot ridings as a grid. The easiest way to do this is using `plot_election()`:

```{r}
plot_election(2015:2019)
```

If you'd like to tweak this plot, you'll have to recreate it using `election_data()`, `layout_province_grid`, and some trickery with `ggplot2::facet_grid()`. The principle is that `layout_province_grid` contains a `geom_x` and `geom_y` column that are a quasi-geographical representation of each riding, grouped by province, and that `ggplot2::geom_tile()` can be used to display this. Because the x and y axes aren't important here, we can use `ggplot2::theme_void()` to hide them.

```{r}
election_results(2015:2019, results = "Elected") %>% 
  select(-province) %>% 
  left_join(layout_province_grid, by = c("election_date", "riding")) %>% 
  mutate(
    party = party %>% fct_infreq(),
    province = fct_rev(province)
  ) %>% 
  ggplot(aes(geom_x, geom_y, fill = party)) +
  geom_tile() +
  scale_fill_party() +
  scale_y_reverse() +
  facet_grid(
    vars(province), 
    vars(year(election_date)),
    scales = "free", 
    space = "free",
    switch = "y"
  ) +
  theme_void() +
  theme(
    strip.text.y = element_text(angle = 180, hjust = 1),
    legend.position = "bottom"
  ) +
  labs(fill = NULL)
```

Even though this representation is far superior to a map, this package wouldn't be complete if it coouldn't plot a map. The e3asiest way to do this is using the `plot_election_map()` function.

```{r}
plot_election_map(2015:2019)
```

If you'd like a different version of this, the `boundaries` dataset within this package contains the polygons for each riding (in [sf](https://r-spatial.github.io/sf/) format). You can join this to the `results` dataset by `c("election_date", "riding"))`.

```{r test-map}
election_results(2015:2019, results = "Elected") %>%
  left_join(boundaries, by = c("election_date", "riding")) %>%
  ggplot(aes(fill = party, geometry = boundary))  +
  geom_sf(size = 0.15) +
  theme_void() +
  scale_fill_party() +
  facet_wrap(vars(year(election_date))) +
  theme(legend.position = "bottom") +
  labs(fill = NULL)
```
